Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-17 paper/manuscript.Rmd
respond to JGH comment
edits

<!-- add numbers in here -->
In total, these XX swimming events occurred throughout the ice-free season, with
a peak in November. On average, caribou swam between islands approximately ever
XX days, but some individuals swam up to XX times per week/month/whatever.
<!--Is it better
to just state our results here, or give this sort of teaser? Initially I had "86
events by 3 individuals" here, but then we ignore that for a couple paragraphs
of context before getting back into it... which seemed weird-->

@@

Forage limitation and over-grazing is a major concern for caribou populations
[@Schaefer2016] and could explain movement between islands. Conspecific
competition for resources driven by reduced foraging opportunities can decrease
reproductive success of adult females [@Schaefer2016].Forage scarcity has been
proposed as a potential reason for caribou moving between arctic islands on the
sea-ice in winter [@Miller1977]. Caribou were introduced to Fogo Island (n = 26
animals between 1964–67) and Change Island (n = 5 animals in 1964) from
Newfoundland as part of a series of translocations and introductions throughout
the province (Bergerud & Mercer 1989). It is unknown when caribou began
colonizing nearby islands, but given that some are small, over-grazing by newly
arrived caribou may rapidly deplete forage [@Bergerud1990]. We posit that some caribou swim to nearby islands and forage

Without sea ice,
movement from the mainland to an island, or between islands, requires caribou to
assess the trade-off associated with swimming.

This trade-off can be understood as density-dependent habitat selection governed
by the Ideal Free Distribution [@Morris1987; @Bradbury2015]. Ideal Free
Distribution theory predicts a fitness equilibrium: when the density in a given
habitat patch has exceeded the optimum for fitness within that patch, animals
should relocate and settle new habitat patches, so that fitness is equal across
all patches [@Bradbury2015]. Islands act as discrete habitat patches in
this case, but the costs associated with swimming create a trade-off when moving
between patches. For caribou in the Fogo Island archipelago, the role of forage
limitation in animal habitat selection patterns remains unknown,
but swimming between islands may in part be explained by forage limitation and
conspecific competition.

>>>>

<!-- add numbers in here -->
In total, these XX swimming events occurred throughout the ice-free season, with
a peak in November. On average, caribou swam between islands approximately ever
XX days, but some individuals swam up to XX times per week/month/whatever.
<!--Is it better
to just state our results here, or give this sort of teaser? Initially I had "86
events by 3 individuals" here, but then we ignore that for a couple paragraphs
of context before getting back into it... which seemed weird
QW: I agree. I've added some Fogo context-->

@@

Forage limitation and over-grazing is a major concern for caribou populations
[@Schaefer2016] and could explain movement between islands. Conspecific
competition for resources driven by reduced foraging opportunities can decrease
reproductive success of adult females [@Schaefer2016].Forage scarcity has been
proposed as a potential reason for caribou moving between arctic islands on the
sea-ice in winter [@Miller1977]. Caribou were introduced to Fogo Island (n = 26
animals between 1964–67) and Change Island (n = 5 animals in 1964) from
Newfoundland as part of a series of translocations and introductions throughout
the province (Bergerud & Mercer 1989). It is unknown when caribou began
colonizing nearby islands, but given that some are small, over-grazing by newly
arrived caribou may rapidly deplete forage [@Bergerud1990]. Although
unconfirmed, it is also possible that Fogo Island has experienced forage
depletion since the introduction of caribou in the 1960s. More likely, however,
is fine-scale competition for foraging resources that drives caribou to swim to
new islands.

Movement between islands requires caribou to assess the trade-off associated
with swimming. This trade-off can be understood as density-dependent habitat
selection governed by the Ideal Free Distribution [@Morris1987; @Bradbury2015].
Ideal Free Distribution theory predicts a fitness equilibrium: when the density
in a given habitat patch has exceeded the optimum for fitness within that patch,
animals should relocate and settle new habitat patches, so that fitness is equal
across all patches [@Bradbury2015]. Islands act as discrete habitat patches in
this case, but the costs associated with swimming create a trade-off when moving
between patches. For caribou in the Fogo Island archipelago, the role of forage
limitation in animal habitat selection patterns remains unknown, but swimming
between islands may in part be explained by forage limitation and conspecific
competition.
